---
output:
  pdf_document: default
  html_document: default
---

title: "Permanova_Zero-impact"
author: "Laure Van den Bulcke & Isolde Cornelis"
date: "3/25/2020"
output: html_document

------
INPUT 
------

Data inlezen. 
Ons experimental design bestaat uit 2 niveaus:
- Impact
- Zone

LET OP! 
We kunnen hierdoor kijken naar het relatief belang van de Zone en de Impact

Daarnaast is het belangrijk om in rekening te houden dat de resultaten maar gebaseerd zijn op 2 Impacts en 3 Zone (dus geen weergave van BDNS), besluiten mag je dus enkel nemen op deze Impacts en Zone en niet veralgemenen. Voorzichtig zijn in verwoording van de resultaten!


Beta diversity:  
```{r Read data}
if(.Platform$OS.type == "unix"){home<-"/home/"} else{
  home<-"//192.168.236.131/"}

# make paths
proj.path <- here("/home/genomics/icornelis/02_ZEROimpact/02_COI/NJ2021")

#1 read data
#setwd(paste0("home/genomics/icornelis/Demultiplex_Liu"))
#upload data
table_unrarefied <- readxl::read_excel(paste0(proj.path,"/OWFvsCoastal_concatenated/results_microDecon/table_unrarefied_concatenated_FullTaxonomicAssignment_clean.xlsx"))
env <- read.csv(paste0(proj.path,"/Step5_Statistics/environmental_data.csv"),  header=TRUE, sep=";") 
table_unrarefied <- data.frame((table_unrarefied), check.names = FALSE)

#remove samples from environmental data that are absent from the rarefied data
keep_samples <- c(colnames(table_unrarefied))
env_unrarefied <- env[env$Niskin.sample %in% keep_samples,]
colorder <- c(env_unrarefied$Niskin.sample)

#select ASVs that are assigned to Animalia
table_unrarefied_AnimaliaASVs <- as.data.frame(table_unrarefied[table_unrarefied$Kingdom %in% c("Animalia"),])
table_unrarefied_AnimaliaASVs <-as.data.frame(table_unrarefied_AnimaliaASVs[!table_unrarefied_AnimaliaASVs$Species %in% "NA",])
table_unrarefied_AnimaliaASVs <- as.data.frame(table_unrarefied_AnimaliaASVs[!table_unrarefied_AnimaliaASVs$Phylum %in% "Chordata",])
table_unrarefied_AnimaliaASVs[is.na(table_unrarefied_AnimaliaASVs)] <- 0

#rm(names_DNA, names_DNA2)

#3: sort by species
#Set the tax names as rownames
taxo <- "Species"
merged_data_unrarefied_Animalia <- aggregate(table_unrarefied_AnimaliaASVs[,1:(ncol(table_unrarefied_AnimaliaASVs)-11)], by= list(as.factor(table_unrarefied_AnimaliaASVs[,taxo])),FUN=sum)
rownames(merged_data_unrarefied_Animalia) <- merged_data_unrarefied_Animalia$Group.1
merged_data_unrarefied_Animalia$Group.1 <- NULL
merged_data_unrarefied_Animalia <- merged_data_unrarefied_Animalia[,colorder]

#(rm(table_unrarefied_Animalia, merged_data_unrarefied_Animalia, taxo))
```


De nodige dataframes worden gemaakt uit de ingelezen data:
1) informatie over de soorten: rijen= stalen, kolommen= soortsnamen

```{r Needed dataframes, include=FALSE}
library(vegan)
#data_soorten <- as.data.frame(t(table_unrarefied_AnimaliaASVs[,1:(ncol(table_unrarefied_AnimaliaASVs)-11)]))
data_soorten <- as.data.frame(t(merged_data_unrarefied_Animalia))
empty_rows <- c(rownames(data_soorten[rowSums(data_soorten) == 0, ]))
data_soorten <- data_soorten[!rownames(data_soorten) %in% empty_rows,]
#data_soorten <- data_soorten[1:51,]  #niet kijken naar pooled seq

keep_samples <- c(rownames(data_soorten))
env <- env[env$Niskin.sample %in% keep_samples,]

data_soorten <- data_soorten %>% filter(!grepl("neg", rownames(data_soorten)))
env <- env %>% filter(!grepl("neg", env$Niskin.sample))

replicaten_name <- rownames(data_soorten)

rownames(data_soorten) <- env$Niskin.sample
data_replicaten <- as.data.frame(rownames(data_soorten))
colnames(data_replicaten) <-"V1"
data_replicaten$Impact <- env$Location
data_replicaten$Biol_replicate <- env$Biol_replicate
data_replicaten$Impact_biol <- paste(data_replicaten$Impact,data_replicaten$Biol_replicate,sep="_")
data_replicaten$Environment <- env$Environment
data_replicaten$Zone <- env$Zone

data_replicaten_OWF <- data_replicaten %>% filter(grepl("_OWF", data_replicaten$Environment))
data_soorten_OWF <- data_soorten[data_replicaten_OWF$V1,]

data_replicaten$Impact <- as.factor(data_replicaten$Impact)
data_replicaten$Impact_biol <- as.factor(data_replicaten$Impact_biol)
data_replicaten$Biol_replicate <- as.factor(data_replicaten$Biol_replicate)
data_replicaten$Environment <- as.factor(data_replicaten$Environment)
data_replicaten$Zone <- as.factor(data_replicaten$Zone)
data_replicaten$Env_Zon <- paste(data_replicaten$Environment,data_replicaten$Zone,sep="-")

data_replicaten_OWF$Impact <- as.factor(data_replicaten_OWF$Impact)
data_replicaten_OWF$Impact_biol <- as.factor(data_replicaten_OWF$Impact_biol)
data_replicaten_OWF$Biol_replicate <- as.factor(data_replicaten_OWF$Biol_replicate)
data_replicaten_OWF$Environmen_ <- as.factor(data_replicaten_OWF$Environment)
data_replicaten_OWF$Zone <- as.factor(data_replicaten_OWF$Zone)
data_replicaten_OWF$Env_Zon <- paste(data_replicaten_OWF$Environment,data_replicaten_OWF$Zone,sep="-")

#data_replicaten_nor <- data_replicaten[!data_replicaten$V1 %in% remove_outliner_Coast,]

  
#diversity_column <- c()
#for (i in 1:nrow(data_replicaten)){
 # Impact <- data_replicaten$Impact[i]
 # if (Impact == "120") {diversity <- "HIGH"}
 # if (Impact == "840") {diversity <- "MEDIUM"}
 # if (Impact == "ZVL") {diversity <- "LOW"}
 # diversity_column <- append(diversity_column, diversity)
#}
#data_replicaten$diversity <- diversity_column
#data_replicaten$diversity <- as.factor(data_replicaten$diversity)


library("xlsx")
write.xlsx(data_soorten, paste0(proj.path,"/Step5_Statistics/data_soorten.xlsx"), sheetName = "Sheet1", col.names = TRUE, row.names = TRUE, append = FALSE)
write.xlsx(data_replicaten, paste0(proj.path,"/Step5_Statistics/data_replicaten.xlsx"), sheetName = "Sheet1", col.names = TRUE, row.names = FALSE, append = FALSE)

```



Een Dissimilarity index wordt berekend. Hier Bray-Curtis

```{R Dissimilarity index }
# JACCARD

#dissimilarity index = jaccard (prensence/absence)
dist_soorten_jaccard <- vegdist(data_soorten, method = "jaccard",binary=TRUE) 

# BRAY
range(data_soorten)

dist_soorten_bray <- vegdist(data_soorten, method = "bray") 
dist_soorten_OWF_bray <- vegdist(data_soorten_OWF, method ="bray")


```

Ons experimental design bestaat uit 2 niveaus:
- Zone
- Impact

H0: „the centroids of the groups, as defined in the space of the chosen resemblance measure, are equivalent for all groups.“. Thus, if H0 were true, any observed differences among the centroids in a given set of data will be similar in size to what would be obtained under random allocation of individual sample units to the groups (i.e., under permuration).


```{R Permanova}
#log root transformation
#data_soorten_transformed <- decostand(data_soorten, method="log")
#data_soorten_transformed <- decostand(data_soorten, method="pa")
#data_soorten_transformed <- decostand(data_soorten, method="hellinger")
data_soorten_transformed <- decostand(decostand(data_soorten, method="tot"), method="max")
dist_soorten_transformed_bray <- vegdist(data_soorten_transformed, method = "bray", binary = F) 

OWF <- c(rownames(data_soorten_OWF))
data_soorten_OWF_transformed <- data_soorten_transformed[rownames(data_soorten_transformed) %in% OWF,]
dist_soorten_OWF_transformed_bray <- vegdist(data_soorten_OWF_transformed, method = "bray", binary = F) 

Impact_p <- data_replicaten$Environment
Zone_p <- data_replicaten$Zone
Impact_OWF_p <- data_replicaten_OWF$Environment
Zone_OWF_p <- data_replicaten_OWF$Zone
Zone_Impact_OWF_p <- data_replicaten_OWF$Env_Zon


Impact_biol_repl_p <- data_replicaten$Impact_biol
Biol_repl_p <- data_replicaten$Biol_replicate
Biol_repl_OWF_p <- data_replicaten_OWF$Biol_replicate


#ASSUMPTIONS
library(pairwiseAdonis)
pairwise.adonis(dist_soorten_transformed_bray, Impact_p, p.adjust.m="bonferroni", reduce=NULL, perm=9999)
pairwise.adonis(dist_soorten_transformed_bray, Zone_p, p.adjust.m="bonferroni", reduce=NULL, perm=9999)
pairwise.adonis(dist_soorten_OWF_transformed_bray, Impact_OWF_p, p.adjust.m="bonferroni", reduce=NULL, perm=9999)
pairwise.adonis(dist_soorten_OWF_transformed_bray, Zone_OWF_p, p.adjust.m="bonferroni", reduce=NULL, perm=9999)
pairwise.adonis(dist_soorten_OWF_transformed_bray, Zone_Impact_OWF_p, p.adjust.m="bonferroni", reduce=NULL, perm=9999)

library(pairwiseAdonis)
pairwise.adonis(dist_soorten_bray, Impact_p, p.adjust.m="bonferroni", reduce=NULL, perm=9999)
pairwise.adonis(dist_soorten_bray, Zone_p, p.adjust.m="bonferroni", reduce=NULL, perm=9999)
pairwise.adonis(dist_soorten_OWF_bray, Impact_OWF_p, p.adjust.m="bonferroni", reduce=NULL, perm=9999)
pairwise.adonis(dist_soorten_OWF_bray, Zone_OWF_p, p.adjust.m="bonferroni", reduce=NULL, perm=9999)
pairwise.adonis(dist_soorten_OWF_bray, data_replicaten_OWF$Env_Zon, p.adjust.m="bonferroni", reduce=NULL, perm=9999)

```
Permanova significant voor Impact en biologische replicaten, niet voor DNA replicaten (residuals).

A significant Permanova means one of three things. 1) There is a difference in the location of the samples (i.e. the average community composition), 2) There is a difference in the dispersion of the samples (i.e. the variability in the community composition), or 3) There is a difference in both the location and the dispersion.
So, if you get a significant Permanova you'll want to distinguish between the three options. That of course is why you need to run the permdisp. If you get a non-significant Permdisp you can conclude the first option above is the correct one. If you get a significant Permdisp then it is either the second or third option (that is there is definitely a difference in dispersion and maybe a difference in location). There is no fool-proof way to distinguish between these two, but looking at an MDS plot of the data will hopefully help you do so (also see note 2 below).


ASSUMPTIONS

In the case of a one-way analysis, the PERMANOVA test using
permutations assumes only that the samples are exchangeable under a true null hypothesis

PERMANOVA works with the assumption that dispersion of the data in your samples is the ~same among each other, so before running PERMANOVA, you must run the betadisper->permutest to know if the dispersions are the same. For that to be true, the permutest has to have a non-significant p-value. Knowing the previous, then you can run the PERMANOVA test, otherwise your interpretations will be wrong
=> dispersion testen met permdisp


```{R, PERMANOVA Bray}

# Advanced applied statistieks

pmv_test_bray <- adonis2(dist_soorten_bray ~   Zone_p, data=data_replicaten, permutations=9999)
pmv_test_bray
pmv_OWF_bray <- adonis2(dist_soorten_OWF_bray ~ Zone_OWF_p * Impact_OWF_p * Zone_OWF_p/Impact_OWF_p, data=data_replicaten_OWF, permutations = 9999)
pmv_OWF_bray

#ASSUMPTIONS

#factor Zone
Zone_betadisper_j <- betadisper(dist_soorten_bray, Zone_p, type="centroid")
Zone_permdisp_j <- permutest(Zone_betadisper_j, permutations=9999)
Zone_permdisp_j #SIGN
plot(Zone_betadisper_j, hull=FALSE, ellipse=TRUE)

#factor Zone OWF
OWF_betadisper_j <- betadisper(dist_soorten_OWF_bray, Zone_Impact_OWF_p, type="centroid")
OWF_permdisp_j <- permutest(OWF_betadisper_j, permutations=9999)
OWF_permdisp_j # SIGN
plot(OWF_betadisper_j, hull=FALSE, ellipse=TRUE)

#factor location
#Impact_betadisper_j <- betadisper(dist_soorten_bray, Impact_p, type="centroid")
#Impact_permdisp_j <- permutest(Impact_betadisper_j, permutations=9999)
#Impact_permdisp_j #SIGN
#plot(Impact_betadisper_j, hull=FALSE, ellipse=TRUE)

#log root transformation
#pmv_test_bray_transformed <- adonis2(dist_soorten_transformed_bray ~   Impact_p, data=data_replicaten, permutations=9999)
#pmv_test_bray_transformed
pmv_test_bray_Zone_transformed <- adonis2(dist_soorten_transformed_bray ~   Zone_p, data=data_replicaten, permutations=9999)
pmv_test_bray_Zone_transformed

pmv_OWF_bray_transformed <- adonis2(dist_soorten_OWF_transformed_bray~ Zone_OWF_p * Impact_OWF_p * Zone_OWF_p/Impact_OWF_p, data=data_replicaten_OWF, permutations = 9999)
pmv_OWF_bray_transformed

#factor Zone
transformed_Zone_betadisper_j <- betadisper(dist_soorten_transformed_bray, Zone_p, type="centroid")
transformed_Zone_permdisp_j <- permutest(transformed_Zone_betadisper_j, permutations=9999)
transformed_Zone_permdisp_j #SIGN
plot(transformed_Zone_betadisper_j, hull=FALSE, ellipse=TRUE)

#factor Zone OWF
transformed_OWF_betadisper_j <- betadisper(dist_soorten_OWF_transformed_bray, Zone_Impact_OWF_p, type="centroid")
transformed_OWF_betadisper_zone <- betadisper(dist_soorten_OWF_transformed_bray, Zone_OWF_p, type="centroid")
transformed_OWF_betadisper_impact <- betadisper(dist_soorten_OWF_transformed_bray, Impact_OWF_p, type="centroid")
transformed_OWF_permdisp_j <- permutest(transformed_OWF_betadisper_j, permutations=9999)
transformed_OWF_permdisp_z <- permutest(transformed_OWF_betadisper_zone, permutations=9999)
transformed_OWF_permdisp_i <- permutest(transformed_OWF_betadisper_impact, permutations=9999)
transformed_OWF_permdisp_j
transformed_OWF_permdisp_z
transformed_OWF_permdisp_i#SIGN
plot(transformed_OWF_betadisper_j, hull=FALSE, ellipse=TRUE, label = F)
plot(transformed_OWF_betadisper_zone, hull=FALSE, ellipse=TRUE, label = T)
plot(transformed_OWF_betadisper_impact, hull=FALSE, ellipse=TRUE, label = T)

```
